Supplementary Materials for: Empowering the use of community driven preprints in ecology and evolution

### 2024-03-22 supp/supp.qmd
Just updating README with where we are at and adding analysis for questions in supp

# Load packages
#install.packages("pacman")
pacman::p_load(tidyverse, here, patchwork, readxl, googlesheets4, janitor, maps, forcats,RColorBrewer, flextable, ggimage, gt, performance, ggthemes, mapproj)

@@

tbl_most <- most_use %>% filter(!is.na(valid_percent)) %>% mutate(Percentage = valid_percent*100) %>% select(country, Percentage)  %>% rename("Country" = country)

@@

>>>>

# Load packages
#install.packages("pacman")
pacman::p_load(tidyverse, here, patchwork, readxl, MASS, googlesheets4, janitor, maps, forcats,RColorBrewer, flextable, ggrepel, ggimage, gt, performance, ggthemes, mapproj)

@@

tbl_most <- most_use %>% filter(!is.na(valid_percent)) %>% mutate(Percentage = valid_percent*100) %>% dplyr::select(country, Percentage)  %>% rename("Country" = country)

@@

## Taxa studied in preprints

```{r, fig-taxa}
#| label: fig-taxa
#| fig-cap: Taxa studied in preprints on *EcoEvoRxiv*.
#| 
cols <- c("plants", "algi", "fungi", "microorganisms", "invertebrates", "vertebrates")

# Summary of the data
sum_taxa <- data2  %>% dplyr::select(all_of(cols))  %>% apply(., 2, sum, na.rm = TRUE)  %>% data.frame()   %>% rownames_to_column(var = "taxa")  %>% rename("n" = ".")  %>% mutate(perc = n/sum(n)*100, labels = paste0(round(perc, 1), "%"))

# Use data to create a pir chart and clean up colour palette

pie <- ggplot(sum_taxa, aes(x = "", y = perc, fill = taxa)) +
  geom_col(color = "black") +
  geom_label_repel(aes(label = labels),
            position = position_stack(vjust = 0.6),
           show.legend = FALSE) +
  scale_fill_viridis_d() +
  coord_polar(theta = "y") + 
  theme_void() + labs(fill = "Taxa")
pie
```

Preprints were published on all major taxomic groups, with the most common being plants (`r sum_taxa$perc[1]`%), vertebrates (`r sum_taxa$perc[6]`%), and invertebrates (`r sum_taxa$perc[5]`%) (@fig-taxa).

## Preprint types

```{r, fig-types}
#| label: fig-types
#| fig-cap: Types of papers posted on *EcoEvoRxiv*.

# Summary of the data
sum_types <- data2  %>% tabyl(type_of_preprint)  %>% rename("type" = type_of_preprint, "n" = n)  %>% mutate(perc = n/sum(n)*100, labels = paste0(round(perc, 1), "%"))

pie <- ggplot(sum_types, aes(x = "", y = perc, fill = type)) +
  geom_col(color = "black") +
  geom_label_repel(aes(label = labels),
            position = position_stack(vjust = 0.6),
           show.legend = FALSE) +
  scale_fill_viridis_d() +
  coord_polar(theta = "y") + 
  theme_void() + labs(fill = "Article Type")
pie
```

Preprints on *EcoEvoRxiv* were diverse in nature, with research articles (`r sum_types[sum_types$type == "research article",]$perc`%) being the most common, followed by reviews and meta-analyses (`r sum_types[sum_types$type == "reviews and meta-analyses",]$perc`%) and opinion papers (`r sum_types[sum_types$type == "opinion",]$perc`%) (@fig-types).

## How does career stage and gender impact preprint use?

```{r, fig-career}
#| label: fig-career
#| fig-cap: Year of first publication for submitting authors of preprints on *EcoEvoRxiv*.

ggplot(data2, aes(x = submitting_author_first_publication_year)) +
  geom_histogram(binwidth = 1, fill = "grey", color = "black") +
  labs(x = "Year of submitting authors first publication", y = "Number of preprints") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

count_dat  <- data2  %>% group_by(submitting_author_first_publication_year)  %>% summarise(n = n())

model <- glm.nb(n ~ submitting_author_first_publication_year, data = count_dat)
res <- summary(model)$coefficients
```

We were interested in getting an understanding of whether career stage is related to preprint use. We used the first publication year of the submitting author as a proxy for career stage. We predicted that early career researchers would be more open to using preprints in Ecology and Evolution.

Interestingly, there was strong evidence that the number of preprints posted decreases with later career stages (negative binomial glm: year slope = `r res[2, "Estimate"]`, SE: `r res[2, "Std. Error"]`, *p* < 0.001, N = `r nrow(count_dat[complete.cases(count_dat),])`). The majority of preprints were submitted by authors who published their first paper in the last 10 years (@fig-career), with the median year of first publication of submitting authors being `r median(data2$submitting_author_first_publication_year, na.rm = TRUE)` (mean = `r round(mean(data2$submitting_author_first_publication_year, na.rm = TRUE),0)`; SD = `r sd(data2$submitting_author_first_publication_year, na.rm = TRUE)`).
